String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-10-20 2023-1-19/crv/sections/conclusion.tex

This work proposes an efficient automata-theoretical approach for solving string constraints with regex-counting and string-length. The approach is based on encoding counting operators in regular expressions by cost registers symbolically instead of unfolding them explicitly. Moreover, this paper proposes automata-size reduction techniques to improve performance further. Finally, we use two benchmark suites comprising 48,843 instances in total to evaluate the performance of our approach. The experimental results show that our approach can solve more instances than the state-of-the-art best solvers, at a comparable or faster speed, especially when the counting and length bounds are large. 
For the future work, we plan to investigate how the symbolic approach can be extended to reason about nested counting operators.

>>>>

This work proposed an efficient automata-theoretical approach for solving string constraints with regex-counting and string-length. The approach is based on encoding counting operators in regular expressions by cost registers symbolically instead of unfolding them explicitly. Moreover, this work proposed automata-size reduction techniques to improve performance further. Finally, we used two benchmark suites comprising 48,843 instances in total to evaluate the performance of our approach. The experimental results show that our approach can solve more instances than the state-of-the-art best solvers, at a comparable or faster speed, especially when the counting and length bounds are large. 
For the future work, we plan to investigate how the symbolic approach can be extended to reason about nested counting operators.

### 2023-10-20 2023-1-19/crv/sections/overview.tex

First, we construct an CEFA for the regular expression $(\Sigma \setminus a)^{\{1, 60\}} (\Sigma \setminus b)^{\{1, 60\}} (\Sigma \setminus c)^{\{0, 60\}}$. Three registers are introduced, say $r_1, r_2, r_3$, to represent the three counting operators;  
the nondeterministic finite automaton (NFA) for $(\Sigma \setminus a)^* (\Sigma \setminus b)^* (\Sigma \setminus c)^*$ is constructed; the updates of registers are added to the transitions of the NFA; the counting bounds are specified by the accepting condition $1 \le r_1 \le 60 \wedge 1 \le r_2 \le 60 \wedge 0 \le r_3 \le 60$, resulting in a CEFA $\aut_1$ illustrated in Figure~\ref{fig:overview}(a). $r_1++$ means that we increment the value of $r_1$ by one after running the transition.
A string $w$ is accepted by $\aut_1$ if, when reading the characters in $w$, $\aut_1$ applies the transitions to update the state and the values of registers, reaching a final state $q$ in the end, and the resulting values of the three registers, say $v_1,v_2,v_3$, satisfy the accepting condition. In addition, we construct other two CEFAs $\aut_2$ for $\Sigma^* c^+$ (see Figure~\ref{fig:overview}(b)) and $\aut_3$ for string length function (see Figure~\ref{fig:overview}(c)). In $\aut_3$, a register $r_4$ is used to denote the length of strings and the accepting condition is $\ltrue$ (See Section~\ref{subsec:regex2cefa} for more details about the construction of CEFA.) Note that we represent the counting operators symbolically by registers instead of unfolding them explicitly.

>>>>

At first, we construct a CEFA for the regular expression $(\Sigma \setminus a)^{\{1, 60\}} (\Sigma \setminus b)^{\{1, 60\}} (\Sigma \setminus c)^{\{0, 60\}}$. Three registers are introduced, say $r_1, r_2, r_3$, to represent the three counting operators;  
the nondeterministic finite automaton (NFA) for $(\Sigma \setminus a)^* (\Sigma \setminus b)^* (\Sigma \setminus c)^*$ is constructed; the updates of registers are added to the transitions of the NFA; the counting bounds are specified by the accepting condition $1 \le r_1 \le 60 \wedge 1 \le r_2 \le 60 \wedge 0 \le r_3 \le 60$, resulting in a CEFA $\aut_1$ illustrated in Figure~\ref{fig:overview}(a). $r_1++$ means that we increment the value of $r_1$ by one after running the transition.
A string $w$ is accepted by $\aut_1$ if, when reading the characters in $w$, $\aut_1$ applies the transitions to update the state and the values of registers, reaching a final state $q$ in the end, and the resulting values of the three registers, say $v_1,v_2,v_3$, satisfy the accepting condition. In addition, we construct other two CEFAs $\aut_2$ for $\Sigma^* c^+$ (see Figure~\ref{fig:overview}(b)) and $\aut_3$ for string length function (see Figure~\ref{fig:overview}(c)). In $\aut_3$, a register $r_4$ is used to denote the length of strings and the accepting condition is $\ltrue$ (See Section~\ref{subsec:regex2cefa} for more details about the construction of CEFA.) Note that we represent the counting operators symbolically by registers instead of unfolding them explicitly.
